---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# antable

<!-- badges: start -->
<!-- badges: end -->

The goal of antable is to simplify the process of performing contingency table-based analyses when starting from a data.frame or tibble object. 

## Installation

You can install the development version of antable from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("dobengjhu/antable")
```

## Example

This is a basic example which shows how to use the `twobytwo()` function to generate a collection of summary statistics related to both the marginal proportions as well as association between the `exposure` and `outcome` variables defined by the user:

```{r, include = FALSE}
set.seed(100)
data <- tibble::tibble(
  sex = c("Male", "Female")[rbinom(100, 1, 0.52) + 1],
  smoking_status = sample(c("Current", "Former", "Never"),
                          100,
                          replace = TRUE,
                          prob = c(0.4, 0.15, 0.45)),
  disease = c("Sick", "Cured")[rbinom(100, 1, 0.2) + 1]
)
```


```{r example}
library(antable)
result <- twobytwo(study_tbl = data,
                   outcome = "disease",
                   exposure = "sex")
summary(result)
```

Sometimes variable values are not ordered in a manner that aligns with the research question of interest. For example, in the scenario above, R will automatically treat `Male` as the exposed group and `Sick` as the outcome of interest because they are each the "highest" value of the respective variables. If we were actually interested in treat females as the exposed group and patients who were cured as cases, we change the manner in which we define `exposure` and `outcome` and provide explicit definition using the general form `"var_name == var_level"`:

```{r}
result <- twobytwo(study_tbl = data,
                   outcome = "disease == 'Cured'",
                   exposure = "sex == 'Female'")
summary(result)
```

In cases where the user is solely interested in cross-tabulation of `exposure` and `outcome`, the `crosstab()` function allows for this more targeted investigation. Furthermore, unlike `twobytwo()` which presumes both `exposure` and `outcome` are binary, `crosstab()` only requires that they are categorical:

```{r}
result_disease <- crosstab(study_tbl = data,
                           outcome = "disease",
                           exposure = "sex")
summary(result_disease)

result_smoking <- crosstab(study_tbl = data,
                           outcome = "smoking_status",
                           exposure = "sex")
summary(result_smoking)
```

Both `twobytwo()` and `crosstab()` can also be used to evaluate a single variable of interest by assigning it to the `outcome` with `exposure = NULL`. Note that `twobytwo()` requires the variable to be binary and will not return any statistics related to measuring association or hypothesis testing, while `crosstab()` can accommodate categorical variables with more than two categories. 

```{r}
result_disease_only <- twobytwo(study_tbl = data,
                                outcome = "disease",
                                exposure = NULL)
summary(result_disease_only)

result_smoking_status_only <- crosstab(study_tbl = data,
                                       outcome = "smoking_status",
                                       exposure = NULL)
summary(result_smoking_status_only)
```

Options for dealing with missing values differ between the two functions:

  - `twobytwo()` excludes all observations that are missing a value for `outcome` or `exposure`
  - `crosstab()` gives users the option of excluding observations that are missing a value for `outcome` or `exposure`, or treating the missingness as an additional category for both variables. 
  
```{r}
data_na <- data
data_na$sex[sample(1:100, 10)] <- NA
data_na$disease[sample(1:100, 10)] <- NA
```
  
```{r}
dplyr::summarise(data_na, 
                 across(c(sex, disease), ~ sum(is.na(.))))

dplyr::filter(data_na,
              is.na(sex) | is.na(disease))
```
  
Each variable is missing 10 observations, respectively, but only one observation is missing information for _both_ sex and disease status. 

```{r}
result_disease_twobytwo <- twobytwo(study_tbl = data_na,
                                    outcome = "disease",
                                    exposure = "sex")
summary(result_disease_twobytwo)

result_disease_crosstab <- crosstab(study_tbl = data_na,
                                    outcome = "disease")
summary(result_disease_crosstab)

result_disease_crosstab_missing <- crosstab(study_tbl = data_na,
                                    outcome = "disease",
                                    include_missing = TRUE)
summary(result_disease_crosstab_missing)
```

  
